Zohmg---a large scale data store for aggregated time-series-based data. [DRAFT]

### 2009-05-13 doc/report/msc-report.tex
doc/report/msc-report.tex: * minor corrections and changes to System analysis chapter.

This chapter will discuss and analyze the system that is to be built. First the
function wanted from the system is presented; while discussing the input and
output characteristics, and possible challenges. Secondly, a rough sketch and
discussion of the system's different parts will be presented. Thirdly, the
requirements of the system will be featured.

Throughout this chapter distinction is made between the developer: who develops
the system, and the user: who uses the product which the developer
creates. Although, in reality, they might be the same person.

@@

Last.fm has a central statistics and metrics tool built on a relational
database. It is possible to import analyzed data to this tool; the data is
presented to the user as graphs. Drawbacks of the system include problems with
new data source additions: they can sometime not be loaded because they are to
big. If a user wants to perform a custom query, manual intervention by the
developer is necessary. There is no way for the system to handle custom queries.

@@

Reporting the analyzed is the data is the final step. This is what the user is
interested in: Viewing the analyzed data. The tool at Last.fm presents the user
with a dashboard that has a selection of graphs on primary metrics; it also has
a list of all available graphs. All the graphs are pre-aggregated, no support
exist for custom queries. Enabling the user to query the aggregated data---for
instance viewing the most popular web browser in a country---would open
possibilities to explore the data. During exploration interesting details could
possibly be discovered; details which may be excluded from the pre-rendered
aggregates.

Since the data sources at Last.fm are time-stamped logs, the constraint to only
handle, time-series was brought upon the system. This decision was reached
mainly to keep the implementation simple.

\subsection*{System parts}

The following three components form the outline of the system: importing,
storing, and exporting data.

At Last.fm there are several different facilities which produce logs, among
them: web servers, scrobbling\footnote{Last.fm users can submit what music they
listen to, this is called scrobbling.}, radio etc. This data reside on a
commodity hardware cluster used for parallel computation. The system needs to be
able to easily import data from this source.

@@

% export.
Due to the several different data sources, there exist several different types
of applications. Therefor the system does not implement any user interface, this
is left for the developer to implement.

@@

The four key requirements of the system are: ease of use, speed, scalability,
and being able to fit in the existing software and hardware eco-system.

@@

Another technique is parallelizing, using a cluster of commodity hardware
machines. Using commodity hardware for computation significantly reduces the
cost, since commodity hardware is cheap; thusly preferred in most situations.

% scaling.
In order for a computational setup, cluster or otherwise, to be efficient it
needs to scale well, this means adding more hardware should add as close as
possible to the hardware's potential addition to the computational power.

@@

The system function part discussed the system's goals and parts. The goals are
to have a system which enables the developer to easily import and analyze new
data sources and to be fast enough to serve the analyzed data at mouse-click
rate to the user. The system's three core parts---importing, storing, and
extracting data---have also been discussed.

Furthermore the requirements on the system were discussed; the four key
requirements are: ease of use, speed, scaling, and fitting into the existing
hardware and software eco-system.

>>>>

This chapter will discuss and analyze the Zohmg system, that is to be built.

First the application area and the wanted system functionality are presented;
while discussing the input and output characteristics, and possible challenges.
Furthermore, a rough sketch and discussion of the Zohmg systems different
components will be presented. Secondly, the Zohmg system requirements will be
featured.

Throughout this chapter distinction is made between the developer: who develops
the application, and the user: who uses the product which the developer
creates. Although, in reality, they might be the same person.

@@

Last.fm has a central statistics and metrics tool built on a relational
database. It is possible to import analyzed data to this tool; the data is
presented to the user as graphs. Disadvantages of this system include problems
with new data source additions: they can sometimes not be loaded because they
are to big. If a user wants to perform a custom query, manual intervention by
the developer is necessary. There is no way for this system to handle custom
queries.

@@

Reporting the analyzed data is the final step. This is what the user is
interested in: Viewing the analyzed data. The tool at Last.fm presents the user
with a dashboard that has a selection of graphs on primary metrics; it also has
a list of all available graphs. All the graphs are pre-aggregated, no support
exists for custom queries. Enabling the user to query the aggregated data---for
instance viewing the most popular web browser in a country---would open
possibilities to explore the data. During exploration interesting details could
possibly be discovered; details which may be excluded from the pre-rendered
aggregates.

Since the data sources at Last.fm are time-stamped logs, the constraint to only
handle, time-series was brought upon the Zohmg system. This decision was
reached mainly to keep the implementation simple.

\subsection*{System components}

The following three components form the outline of the Zohmg system: importing,
storing, and exporting data.

At Last.fm there are several different facilities which produce logs, among
them: web servers, scrobbling\footnote{Last.fm users can submit what music they
listen to, this is called scrobbling.}, radio streaming etc. This data reside
on a commodity hardware cluster used for parallel computation. The system needs
to be able to easily import data from this source.

@@

% export.
Due to the several different data sources, there exist several different types
of applications. Therefor the Zohmg system does not implement any user
interface, this is left for the developer to implement.

@@

The four key requirements of the Zohmg system are: ease of use, speed,
scalability, and being able to fit in the existing software and hardware
eco-system.

@@

The other technique, parallelizing, uses a cluster of commodity hardware
machines. Using commodity hardware for computation significantly reduces the
cost, since commodity hardware is cheap; thusly preferred in most situations.

% scaling.
In order for a computational setup, cluster or other, to be efficient it needs
to scale well, this means adding more hardware should add as close as possible
to the hardware's potential addition to the computational power.

@@

The system function section discussed the Zohmg system's goal and its
components. The goal is to have a system which enables the developer to easily
import and analyze new data sources and to be fast enough to serve the analyzed
data at mouse-click rate to the user. The system's three core
components---importing, storing, and extracting data---have also been
discussed.

Furthermore the requirements on the Zohmg system were discussed; the four key
requirements are: ease of use, speed, scaling, and fitting into the existing
hardware and software eco-system.
